---
title: "Homework_1"
author: "Warren Geither"
date: "8/29/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

In class we discussed the Binomial distribution. Another discrete distribution is the Poisson distribution. 

a.) Look online and in your book collection to read about the Poisson distribution. Is f(y|lambda) a pdf or     pmf?
    - pmf

b.) What is the expected value E(y)? Prove it.
    - see attached pdf

c.) What is the variance var(y)? Prove it.
    - see attached pdf

d.) The number of mistakes a letter stuffing robot makes in an hour can be considered to be a Poisson     random variable. Let the random variable Y have a Poisson distribution with parameter lambda=2.3.

  - What is the probability that the robot makes 2 mistakes in an hour?
```{r}
dpois(2, lambda = 2.3)
```
 
  - What is the probability the robot makes at least 2 mistakes in an hour?
```{r}
1 - sum(dpois(0:2, lambda = 2.3))
```

  - What is the the probability that the robot makes 2.3 mistakes in an hour?

  - What is the probability that the robot makes 0 mistakes in an hour?

  - What is the probability that the robot makes less than 6 mistakes in an hour?